---
title: "Machine Learning Project"
output: html_notebook
---

Importing Packages : 
--------------------
```{r}
library("FactoMineR")
library("factoextra")
library(mgcv) # library for GAM
library(ggplot2) # for beautiful plots
library(cdata) # data wrangling
library(sigr) # AUC calculation
library(xtable) # to save table in latex format
```

Importing Data : 
---------------

```{r}
X_sm = read.csv('X_train_smote.csv')
y_sm = read.csv('y_train_smote.csv')
X_train = read.csv('x_no_smote.csv')
Y_train = read.csv('y_no_smote.csv')
X_test = read.csv('X_test.csv')
Y_test = read.csv('Y_test.csv')
x_sm_factor_analysis <- X_sm
x_test_factor_analysis <-X_test
```

Factor Analysis of Mixed Data FAMD : 
-----------------------------------


## Data Preperation:

The data type is as follows:

```{r}
str(x_sm_factor_analysis)
```

In order to apply FAMD on my data, the categorical variables that are in numeric form will be converted to string form

```{r}
categorical <- c('Household.income' , 'Home.Ownership','Household.in.urban.rural.area' , 
                 'population.density' , 'Price.of.Gasoline.Affects.Travel'  ,
                 'Travel.is.a.Financial.Burden' , 'Gender' , 'Educational.Attainment' ,
                 'More.than.One.Job' , 'Job.Category' , 'Relationship' ,
                 'Full.Time.or.Part.Time.Worker' , 'Level.of.Physical.Activity', 
                 'Public.Transportation.or.Taxi' , 'Passenger.to.Friend.Family.Member.or.Rental.Car' ,
                 'Bicycle.or.Walk' , 'Race.African' , 'Race.Asian' , 'Race.American.Indian' ,
                 'Race.Native.Hawaiian' , 'Multiple.Response.Race' )
                 
x_sm_factor_analysis[categorical]  <- lapply(x_sm_factor_analysis[categorical] , as.character)
x_test_factor_analysis[categorical]  <- lapply(x_test_factor_analysis[categorical] , as.character)
```

Now after conversion, the data type is as follows:

```{r}
str(x_sm_factor_analysis)
```


## Model Fitting and Analysing


```{r}
res.famd <- FAMD(x_sm_factor_analysis,  ncp = 70, sup.var = NULL, ind.sup = NULL, graph = FALSE)
```

The proportion of variance retained by the different dimentions is the following:

```{r}
eig.val <- get_eigenvalue(res.famd)
eig.val
```

The scree plot is the following :

```{r}
fviz_screeplot(res.famd)
```
## Graph of Variable:

```{r}
var <- get_famd_var(res.famd)
var
```
```{r}
# Plot of variables
fviz_famd_var(res.famd, repel = TRUE)
# Contribution to the first dimension
fviz_contrib(res.famd, "var", axes = 1)
# Contribution to the second dimension
fviz_contrib(res.famd, "var", axes = 2)
```
```{r}
quanti.var <- get_famd_var(res.famd, "quanti.var")
quanti.var 
```


```{r}
fviz_famd_var(res.famd, "quanti.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)
```

Now I will save the components to excel so I can test their predictive performance in jupyter notebook.

```{r}
res_train <- FAMD(x_sm_factor_analysis,  ncp = 70, sup.var = NULL, ind.sup = NULL, graph = FALSE)
res_test <-FAMD(x_test_factor_analysis,  ncp = 70, sup.var = NULL, ind.sup = NULL, graph = FALSE)
x_famd_train <- data.frame(res_train$ind$coord)
x_famd_test <-data.frame(res_test$ind$coord)
```

Now I'll save the results to excel file so I could evaluate their predictive performance
```{r}

write.csv(x_famd_train,'x_famd_train.csv',row.names = FALSE)
write.csv(x_famd_test ,'x_famd_test.csv',row.names = FALSE)
```


Now I will apply some codes that transforms the eigenvalues results to latex table

```{r}
print(xtable(eig.val, type = "latex"), file = "eignvalue.tex")
```

# Logistic Gam Model : 

```{r}

train <- data.frame(y_sm,X_sm)

# formula: “Outcome==1” is to binarize the response variable
form_gam=as.formula(" AFV == 1~s(Household.income,k = 6) +Home.Ownership+s(Count.of.household.members,k = 5)+Household.in.urban.rural.area+s(population.density,k = 5)+s(Price.of.Gasoline.Affects.Travel,k = 4)+s(Travel.is.a.Financial.Burden,k = 4)+s(Age)+Gender+s(Educational.Attainment,k = 4)+More.than.One.Job+s(Job.Category,k = 4)+s(Count.of.Car.Share.Program.Usage,k = 6)+s(Trip.Time.to.Work.in.Minutes)+s(Best.estimate.of.annual.miles)+s(Count.of.household.vehicles,k = 4)+s(count.of.children.in.household,k = 3)+s(Relationship,k = 5)+Full.Time.or.Part.Time.Worker+s(Count.of.Walk.Trips)+s(Count.of.Bike.Trips)+s(Level.of.Physical.Activity,k = 3)+s(Count.of.adult.household.members,k = 3)+s(Public.Transportation.or.Taxi,k = 3)+s(Passenger.to.Friend.Family.Member.or.Rental.Car,k = 3)+s(Bicycle.or.Walk,k = 3)+Race.African+Race.Asian+Race.American.Indian+Race.Native.Hawaiian +Multiple.Response.Race ")
gam_model=gam(form_gam,data=train,family = binomial(link="logit"))
gam_model$converged # check if the algorithm converges
summary(gam_model)
```


```{r}
#plotting non linear components
par(mfrow=c(1,3), mar=c(4.1, 4.1, 2.1, 2.1)) 
plot(gam_model, shade=TRUE, shade.col="lightblue")

```
 Now I will save GAM model results to csv in order to evaluate the prediction in my jupyter file so I could have the same form of plot 
```{r}
prediction_gam <- predict(gam_model, newdata = X_test, type="response")
prediction_dataframe = data.frame(prediction_gam)
write.csv(prediction_dataframe,'prediction_gam.csv',row.names = FALSE)
```

  